Naturalness and Long-Lived Beyond the Standard Model Particles

### 2017-07-31 tex/draft.tex
Add pixel figure, edit llp section

    The mechanism and scale of SUSY breaking is an open question with many competing answers that achieve different theoretical goals and lead to different TeV-scale phenemena. A handful of models with particular approaches to SUSY breaking that are particularly prominant or particularly relevant to naturalness or long-lived particles are gauge-mediated susy breaking (GMSB), anomaly-mediated SUSY breaking (AMSB), split\fxnote{capitalize} SUSY, and R-parity violating SUSY (RPV). The details of these models will be explored as needed throughout this paper.

@@

    On the other hand, models with weakly coupled R-parity violating (RPV) terms produce long-lived but not perfectly stable LSPs. Fig.~\ref{rp_and_rpv_diagrams} shows two possible processes, one where R-parity gaurantees a stable LSP and one where RPV produces a long-lived but stable LSP. A similar situation arises in GMSB models\fxnote{should I describe GMSB?} where the gravitino is the LSP. The coupling between the next-to-LSP (NLSP) and gravitino is inversely proportional to the scale of SUSY breaking, which supresses the NSLP decay rate and can lead to LLPs.\fxnote{cite dimopoulos low energy}

    \noindent \begin{figure}[htbp] \begin{center}
    \includegraphics[width=0.3\textwidth]{images/rp_and_rpv_diagrams}
        \caption{R-parity and RPV diagrams}
    \label{rp_and_rpv_diagrams}
    \end{center} \end{figure}

    LLPs can also arise from particular SUSY mass spectra. Models in the Split\fxnote{capitalize?}  SUSY paradigm, for example, propose that the scalar sparticles are significantly more massive than the gauginos. In these models, the gluino becomes long-lived when its decay to a quark and neutralino is mediated by a highly-virtual squark. \fxnote{cite kilian} \fxnote{mention that split throws out naturalness motivation?}\fxnote{Add diagrams} On the other hand, some AMSB models\fxnote{explain AMSB?} predict the zino\fxnote{no such thing. Maybe I meant bino or general neutralino?} and wino are the LSP and NLSP, respectively, and are nearly degenerate in mass. The wino then becomes long-lived when its decay is suppressed by the lack of available phase space. \fxnote{cite R\&S out of this world}
    \fxnote{Andrew says don't worry about mentioning models that don't come up in the searches, but be prepared to explain the basic ideas behind HV, dark sector and how they result in LLP}

@@

        The pixel tracker, which was upgraded in winter 2016/17, makes up the innermost layer and features approximately \num{124} million silicon pixels, each of which covers an active area of \num{100} by \SI{150}{\micro\meter^2}. The pixels are spread between four barrel layers (BPix) at $\num{2.9}<\mathrm{r}<\SI{16}{cm}$ and 3 endcap disks (FPix) at $\num{29} < \mathrm{z} < \SI{52}{cm}$. The upgraded pixel tracker should provide 4-hit tracking out to $\eta < \num{2.4}$ and improved b-tagging over the original pixel tracker despite the increased pileup expected in 2017.

@@

        This search\fxnote{repetative wording} uses a simplified benchmark model in the split SUSY paradigm where gluinos are kinematically accessible but squarks are not. The relevant process, shown in Fig.~\ref{displaced_process}, is pair-produced gluinos that decay to a quark-\fxnote{how to punctuate?}virtual squark pair. The virtual squark then decays to a quark and the LSP, presumed to be a neutralino\fxnote{have I defined neutralino?}\fxnote{why is this a reasonable assumption?} \fxnote{steal diagram from andrew's preapproval. I can't access it - ask andrew}. The large mass of the intermediate squark suppresses the gluino decay, giving the gluino enough time to form an R-hadron\fxnote{first use. Define here if not in LLP section} that propagates some measureable distance before decaying.

@@

        This analysis is interpreted in the context of two potential LLPs: gluinos and multiply charged massive particles (mchamps). Long-lived gluinos can appear in models such as split\fxnote{capitalize?} SUSY, where the squarks and sleptons are quite heavy, but the gluino is relatively light. After being pair produced, gluinos form R-hadrons\fxnote{have these been defined?} and eventually decay to quark-antiquark pairs along with NLSP neutralinos, which then decay to LSP neutralinos along with muon-antimuon pairs. This process is exactly the same as in Fig.~\ref{displaced_process}, but in this case the gluino has enough time to form R-hadrons\fxnote{double check}. The gluino decay is supressed by the large squark mass, which leads to its appreciable lifetime.

>>>>

    The mechanism and scale of SUSY breaking is an open question with many competing answers that achieve different theoretical goals and lead to different TeV-scale phenemena. A handful of models with particular approaches to SUSY breaking that are particularly prominant or particularly relevant to naturalness or long-lived particles are gauge-mediated susy breaking (GMSB), anomaly-mediated SUSY breaking (AMSB), Split SUSY, and R-parity violating SUSY (RPV). The details of these models will be explored as needed throughout this paper.

@@

    On the other hand, models with weakly coupled R-parity violating (RPV) terms produce long-lived but not perfectly stable LSPs. A similar situation arises in GMSB models\fxnote{should I describe GMSB?} where the gravitino is the LSP. The coupling between the next-to-LSP (NLSP) and gravitino is inversely proportional to the scale of SUSY breaking, which supresses the NSLP decay rate and can lead to LLPs.\fxnote{cite dimopoulos low energy}

    \fxnote{diagram or eqs for R-parity / RPV?}

    LLPs can also arise from particular SUSY mass spectra. Models in the Split SUSY paradigm, for example, propose that the scalar sparticles are significantly more massive than the gauginos. In these models, the gluino becomes long-lived when its decay to a quark and neutralino is mediated by a highly-virtual squark. \fxnote{cite kilian} \fxnote{mention that Split throws out naturalness motivation?} On the other hand, some AMSB models\fxnote{explain AMSB?} predict the winos as the lightest gauginos, which leads to nearly-degenerate LSP and NLSP. The NLSP wino then becomes long-lived when its decay is suppressed by the lack of available phase space. \fxnote{cite R\&S out of this world via disappearing}

    \fxnote{show a few SUSY spectra?}

@@

        The pixel tracker, which was upgraded in winter 2016/17, makes up the innermost layer and features approximately \num{124} million silicon pixels, each of which covers an active area of \num{100} by \SI{150}{\micro\meter^2}. The pixels are spread between four barrel layers (BPix) at $\num{2.9}<\mathrm{r}<\SI{16}{cm}$ and 3 endcap disks (FPix) at $\num{29} < \mathrm{z} < \SI{52}{cm}$. The upgraded pixel tracker, shown in Fig.~\ref{pixel_old_new}, should provide 4-hit tracking out to $\eta < \num{2.4}$ and improved b-tagging over the original pixel tracker despite the increased pileup expected in 2017.

        \noindent \begin{figure}[htbp] \begin{center}
            \includegraphics[width=0.5\textwidth]{images/pixel_old_new}
            \caption{Diagram showing barrel pixel layers before and after the winter 2016/17 upgrade.\fxnote{cite website found through Andrew's citations}}
            \label{cms}
        \end{center} \end{figure}

@@

        This search\fxnote{repetative wording} uses a simplified benchmark model in the Split SUSY paradigm where gluinos are kinematically accessible but squarks are not. The relevant process, shown in Fig.~\ref{displaced_process}, is pair-produced gluinos that decay to a quark-\fxnote{how to punctuate?}virtual squark pair. The virtual squark then decays to a quark and the LSP, presumed to be a neutralino\fxnote{have I defined neutralino?}\fxnote{why is this a reasonable assumption?} \fxnote{steal diagram from andrew's preapproval. I can't access it - ask andrew}. The large mass of the intermediate squark suppresses the gluino decay, giving the gluino enough time to form an R-hadron\fxnote{first use. Define here if not in LLP section} that propagates some measureable distance before decaying.

@@

        This analysis is interpreted in the context of two potential LLPs: gluinos and multiply charged massive particles (mchamps). Long-lived gluinos can appear in models such as Split SUSY, where the squarks and sleptons are quite heavy, but the gluino is relatively light. After being pair produced, gluinos form R-hadrons\fxnote{have these been defined?} and eventually decay to quark-antiquark pairs along with NLSP neutralinos, which then decay to LSP neutralinos along with muon-antimuon pairs. This process is exactly the same as in Fig.~\ref{displaced_process}, but in this case the gluino has enough time to form R-hadrons\fxnote{double check}. The gluino decay is supressed by the large squark mass, which leads to its appreciable lifetime.
